For submission to the Journal of Integrated Pest Management

### 2021-09-07 ofv_fife_2021.Rmd
Addressing comments to authors: Dr. Nathan Walker: My Comments to Authors: Line 40 and throughout the manuscript, not needed hard returns are present and need to be removed. Lines 49-50 Coinfections were seen in each county, but single strains of OFV-Orc were seen in L. muscari – Can coinfections be seen? Can single strains be seen?  No, symptoms can be and coinfections are determined by analysis. The Spanish abstract addition is not warranted and I am confused why you added this at this late stage? The staff editors have suggested “So, my suggestion would be just to assume the other [English]

We describe the first detection of orchid fleck virus (OFV) infecting three unreported hosts: *Liriope muscari*, cv. 'Gigantea' (Decaisne) Bailey, *Ophiopogon intermedius* Don and *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) in Leon and Alachua Counties, FL. Orchid-infecting subgroup (Orc) of OFV infects over 50 plant species belonging to the plant families Orchidaceae, Asparagaceae (Nolinoidaea), and causes citrus leprosis disease in *Citrus* (Rutaceae). The only known vectors of OFV-Orc are the flat mites, *Brevipalpus californicus* (Banks) *sensu lato* (Trombidiformes: Tenuipalpidae). Florida has various plants in the landscape which *Brevipalpus* spp. feed on, which are susceptible to infection by OFV-Orc.

Chlorotic ringspots and flecking were seen affecting Liriopogons (*Liriope* and *Ophiopogon* spp.) in Leon County, FL. Nearby *A. elatior* also appeared chlorotic. Local diagnostics returned negative for common plant pathogens, therefore new samples were sent to the Florida Department of Agriculture and Consumer Services (FDACS) and USDA-ARS for identification.

Two orchid-infecting strains of OFV were detected via combinations of conventional RT-PCR, RT-qPCR, Sanger sequencing and High Throughput Sequencing (HTS). Amplicons shared 98\% nucleotide identity with OFV-Orc1 and OFV-Orc2 RNA2 genome sequences available in NCBI GenBank. Coinfections were seen in each county, but single strains of OFV-Orc were seen in *L. muscari* (Alachua, OFV-Orc2) and *A. elatior* (Leon, OFV-Orc1).

Three potential mite vectors were identified via cryo-scanning electron microscopy (Cryo-SEM): *Brevipalpus californicus* (Banks) sensu lato, *B. obovatus* Donnadieu, and *B. confusus* Baker.

In conclusion, OFV orchid strains are present in northern Florida, representing a risk for susceptible plants in the southeastern US.

@@

Se describe la primera detección del virus de orchid fleck virus (OFV), infectando a tres huéspedes no reportados: *Liriope muscari*, cv. ‘Gigantea’ (Decaisne) Bailey, *Ophiopogon intermedius* Don and *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) para los condados de Leon y Alachua, FL. Los subgrupos de OFV que infectan a las orquídeas (Orc) puedan infectar más de 50 especies de plantas pertenecientes a las familias Orchidaceae, Asparagaceae (Nolinoidaea), e infecta *Citrus* (Rutaceae) cómo la enfermedad de la leprosis de los cítricos. Los únicos vectores de OFV-Orc son los ácaros planos *Brevipalpus californicus* (Banks) *sensu lato* (Trombidiformes: Tenuipalpidae). La Florida tiene varias plantas en el campo ornamental, por lo cúales las especias de *Brevipalpus* se puedan alimentarse, y esas plantas son susceptibles a las infecciones de OFV-Orc. Se observaron manchas anulares cloróticas y salpicaduras en las hojas de las Serpentinas (*Liriope* y *Ophiopogon* spp.) y también se vieron hojas cloróticas en el *A. elatior* adyacente, situado en el condando de Leon, FL. Los diagnósticos del laboratorio local fueron negativos para los patógenos comunes, por lo tanto, se enviaron nuevos ejemplares al Florida Department of Agriculture and Consumer Services (FDACS) y el USDA-ARS para su identificación.
Se detectaron dos cepas del OFV mediante la combinación de RT-PCR convencional, RT-qPCR, secuenciación de Sanger y secuenciación de alto rendimiento. Las ampliaciones compartían una identidad de nucleótidos del 98% con las secuencias del genoma de ARN2 de OFV-Orc1 y OFV-Orc2 disponibles en el NCBI GenBank. Se observaron coinfecciones del virus en cada condado, pero se observaron cepas únicas de OFV-Orc en *L. muscari* (Alachua , OFV-Orc2) y *A. elatior* (Leon , OFV-Orc1).
Se identificaron tres ácaros mediante microscopía electrónica de barrido criogénico (Cryo-SEM) con potencial de ser vectores: *Brevipalpus californicus* (Banks) *sensu lato*, *B. obovatus* Donnadieu y *B. confusus* Baker. En conclusión, OFV está presente en el norte de Florida, lo que representa un riesgo para las plantas susceptibles en el sureste de Estados Unidos.

@@

\pagebreak
Orchid fleck virus (OFV), is the type member for the genus *Dichorhavirus*, family *Rhabdoviridae*. OFV is a bacilliform, nuclear rhabdovirus composed of two segments of single-stranded, negative-sense RNA which infects plants [@Amarasinghe2019; @Dietzgen2014; @Walker2018]. Flat mites from the genus *Brevipalpus* (Trombidiformes: Tenuipalpidae) are the only which have been experimentally shown to transmit dichorhaviruses [@Maeda1998].

OFV-infected plants exhibit various symptoms depending on the infected plant species as well as the strain of the OFV associated with the infection [@Kubo2009a; @Kubo2009b], but symptoms typically appear as chlorotic flecks, which ultimately coalesce into ringspot patterns.

OFV was first described as infecting *Cymbidium* orchids in Japan [@Doi1977]. OFV and OFV-like rhabdoviruses have been reported infecting orchids in Asia, Africa, North America, South America, Europe, and Oceania. The prevalence of OFV and its mite vector is thought to be associated with the movement of infected orchids [@Dietzgen2018a].

OFV naturally infects more than fifty species of Orchidaceae [@Kitajima2010; @Peng2013], some Asparagaceae (Nolinoidaea) [@Mei2016; @Dietzgen2018b], and Rutaceae, where it causes citrus leprosis-like symptoms [@Cook2019; @Velarde2021; @Roy2015; @Roy2020]. Mechanical transmission of OFV is possible under laboratory conditions to some plants belonging to the plant families Chenopodiaceae, Aizoaceae, Fabaceae, and Solanaceae [@Chang1976; @Kondo2003; @Peng2013].

@@

During June 2020, chlorotic ringspot symptoms were observed on Giant Lilyturf *Liriope* spp., cv. 'Gigantea' in a landscape of Leon County, Florida (`r citef("fig_1")`). *Liriope* belong to a group of plants in the family Asparagaceae, subfamily Nolinoidaea, comprised of grass-like monocotyledonous liliod plants native to southeastern Asia [@Chase2009; @Meng2021]. *Liriope* and the closely related *Ophiopogon* (Asparagaceae: Nolinoidaea) are considered the most important ground cover plant in the southeastern United States [@Mcharo2003].

Viral infections of suspected leaf samples were initially tested at the Plant Disease Diagnostic Clinic at the North Florida Research and Education Center (NFREC) in Quincy, FL. All the samples were tested with one step conventional RT-PCR, and were found negative for begomovirus, carlavirus, potyvirus, tospovirus, cucumber mosaic virus and tobacco mosaic virus.

As initial diagnostics were inconclusive, samples were taken of putatively infected plants with ringspot symptoms during July and August of 2020. Leaves were taken from *Liriope* spp. and *Ophiopogon* spp., as well as the *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea), nearby, which appeared sickly and chlorotic (`r citef("fig_2")`). Plant materials were sent to the Florida Department of Agriculture and Consumer Services (FDACS) for identification.

The FDACS determined that the pathogen was OFV using previously published primers and methods to conduct RT-PCR and Sanger sequencing [@Kubo2009a; @Kubo2009b; @RamosGonzalez2015]. OFV Orchid strain 1, (OFV-Orc1), was identified following the methods described in @Kondo2017. Nucleotide sequencing shared 98\% nucleotide identity with the OFV-isolates So (Accession No. AB244418) and Br (Accession No. MK522807), which belong to orchid subgroup I [@Kondo2006; @Kondo2017].

These samples from FDACS were subsequently retested by the USDA-APHIS-PPQ S&T Beltsville laboratory, in conjunction with tests of fresh samples from both Alachua and Leon counties. The USDA used RT-PCR, RT-qPCR, and HTS to reconfirm the presence of OFV. Conventional RT-PCR with Generic R2-Dicho-GF and R2-Dicho-GR primers amplifed ~800 nt of L-gene (RNA2) amplicon [@Roy2020], and OFV-Orc1 and OFV-Orc2 were detected in both *O. intermedius* and *A. elatior* from Leon County [@Kondo2017] (GenBank Accession Numbers: MZ852004, MZ852005 MZ852006, and MZ852007). OFV-Orc1 and OFV-Orc2 share 99% nucleotide sequence identity with their RNA2 genome wheras 90\% nt sequence identity was found between these two reassortment strains.

HTS reaffirmed the presence of OFV-Orc1 and OFV-Orc2 in Leon and Alachua counties (`r citet("tab_1")`). HTS results from Leon County revealed that *L. muscari* were coinfected with both OFV-Orc1 and OFV-Orc2, while *A. elatior* were solely infected with OFV-Orc1. HTS of *L. muscari* from Alachua County revealed infections with OFV-Orc2 (GenBank Accession MZ852006).

After the initial identification by FDACS of OFV-Orc, mite samples were collected from symptomatic Asparagaceae in Leon County. Most mites collected were Tenuipalpid mites (flat mites or false spider mites), a pest of ornamental plants, some of which are known to act as vectors for plant viruses [@Childers2003; @Childers2011].

@@

Mite taxonomy is complicated by cryptic species complexes which occur in many plant-feeding groups of the Acari [@Arthur2011; @Skoracka2010; @Skoracka2013; @Umina1999], including tenuipalpid mites from the genus *Brevipalpus* [@Navia2013]. The commonly used phase-contrast microscopy is insufficient to detect some diagnostic characters for separation of cryptic species, instead best practices recommend the combination of Differential Interference Contrast (DIC) Microscopy and Scanning Electron Microscopy along with molecular methods to separate cryptic species [@Beard2015].

The flat mites collected were initially suspected to belong to *B. californicus* after inspection with phase contrast microscopy. Subsequent observation via DIC microscopy at FDACS agreed with this tentative identification. Unfortunately, the *B. californicus* s.l. species group, *sensu* @Baker1987 is suspected to contain cryptic species [@Childers2011; @Rodrigues2013]. New mite samples were collected from symptomatic liriopogons and *A. elatior* in Leon County and sent to USDA-ARS's Electron and Confocal Microscopy Unit for analysis. Three mite species were recovered and examined under cryo-scanning electron microscopy (Cryo-SEM): *B. californicus* s.l. (`r citef("fig_3")`), *B. obovatus* Donnadieu and *B. confusus* Baker.

The recent report of OFV in the US is thought to be @Ko1985 which describes nuclear inclusions caused by an undescribed bacilliform rhabdovirus in *Brassia* orchids. The significance of this report is their description of the spoke-wheel configurations of the viral particles [@Ko1985], a sign typically associated with OFV infection [@Chang1976]. Unfortunately, this article made no mention of mites or further investigations of the virus. The first report of OFV in the continental US was @Bratsch2015, who confirmed the presence of OFV in *Phalaenopsis* hybrids using Transmission Electron Microscopy of ultrathin sections of plant tissue as well as molecular sequence analysis. They also discuss the association of OFV with *Brevipalpus* mites, but the authors did not make a conclusive species identification beyond suggesting that the mite vector belonged to the *B. californicus* group, referring to @Kondo2003's publication [@Bratsch2015].

Later reports of OFV described OFV infecting a previously undescribed Nolinoidaea hosts in Australia [@Dietzgen2018b; @Mei2016], including *Liriope spicata* (Thunb.) Lour, a different species of liriopogon than those identified from the Florida sites. We are not aware of any reports of OFV infecting liriopogons, *A. elatior* nor other Nolinoidaea in the US. Although @Zheng2013 had mentioned an association between *B. californicus* and *A. elatior*, they never reported symptoms of OFV-Orc in this plant. We believe that our findings indicate the first report of OFV-Orc infecting ornamental Nolinoidaea in Florida, and possibly the US. This publication also marks the first reports of *A. elatior* and *Ophiopogon* spp. as natural hosts of OFV-Orc.

OFV consists of two orchid strains (OFV-Orc1 and OFV-Orc2) and two citrus strains (OFV-Cit1 and OFV-Cit2) [@Roy2020; @Beltran-Beltran2020]. The OFV strains detected in Florida are identical in gene order, content, and genome sequence to the orchid strains of OFV infecting citrus in Hawaii, Mexico, Colombia, and South Africa [@Roy2020; @Beltran-Beltran2020]. Both OFV-Orc1 and OFV-Orc2 infect citrus [@Roy2020], but none of the citrus  strains have been reported from any orchid species. The *Brevipalpus* mites collected from liriopogons and *A. elatior* in Leon County were abundant on OFV-infected plants very near to citrus trees, some plants even surrounding the trunk. *B. californicus* s. l. has been reported as a pest of citrus [@Childers2003] and are often collected from citrus fruits [@Vacante2016; @Vacante2010; @Baker1987; @Baker1949]. The proximity of these mite vectors to citrus raises the question: why these trees are not currently infected with OFV-Orc? It is important to note the uncertainty surrounding the vector for OFV-Orc. There are three mite species which have been recovered from OFV-Orc infected plants: *B. obovatus*, and *B. confusus* and *B. californicus* s.l., but only *B. californicus* has been described as a vector of OFV. Even so, the *B. californicus* which we find on liriopogons and *A. elatior* may not be the same cryptic species as those found on citrus. Transmission of OFV from populations of *B. californicus* liriopogon/*A. elatior* to citrus may be limited by host preferences, vectorial capacity, viral propagation/circulation in the vector, viral acquisition times, or feeding times required for transmission to citrus. Even so, these types of questions require future study to determine the potential of nolinoidaea to citrus transmission.

Best practices for integrated pest management have not been created for controlling *Brevipalpus* mites on these ornamentals, but methods designed to control *Brevipalpus* in other systems may be applicable. The most common method used to control *Bervipalpus* are synthetic acaricides [@Andrade2010; @Andrade2019]. Unfortunately, some acaricides and their residues can harm beneficial predatory mites as well [@Fernandez2017], even at low doses [@Havasi2021], and mixing different chemistries can be detrimental for mite control [@Vechia2018]. In addition, pesticide resistance has been reported in various *Brevipalpus* populations [@Rocha2021; @Alves2000; @Campos2002; @Omoto2000], due to exposure to pesticides used to control other arthropod pests [@Vechia2021]. In addition, predatory mites [@Argolo2020; @Chen2006], entomopathogenic fungi [@RossiZalaf2008; @Magalhaes2005; @Pena2015; @Revynthi2019] have shown promise for controlling other *Brevipalpus* mites. Moreover, it is often possible to integrate different control techniques for improved management, such as combining predatory mites with compatible acaricides and entomopathogenic fungi [@Reddy2001; @Midthassel2016; @Andrade2019].

In conclusion, detecting OFV in Florida represents a concern for horticulturists who grow orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagaceae species which are commonly used in landscaping. Florida is also home to a plethora of native and naturalized orchid species, many of which are threatened, including cultivated *Vanilla* in southern Florida [@Chambers2019] and the famous Ghost Orchid, [*Dendrophylax lindenii* (Lindl.) Benth. ex Rolfe]. Citrus leprosis was present in Florida during the 1860’s and almost eradicated by the mid-1960s [@Childers2003; @Knorr1968a; @Knorr1968b]. An examination of herbarium specimens of Florida citrus found that this historical virus, Citrus leprosis dichorhavirus-N0, is distantly related to the modern isolates of OFV [@Hartung2015; @Kitajima2011a; @Roy2020]. The recent detection of OFV-Orc1 in South Africa [@Cook2019] in *C. sinensis* (Navel and Valencia orange) and OFV-Orc2 in Hawaii [@Velarde2021] in *C. reticulata* (mandarin) and *C. jambhiri* (rough lemon) associated with leprosis-like symptoms highlights the potential threat of different isolates of OFV on citrus, which will be a definite concern to the US multi-billion-dollar citrus industry already impacted by the Huanglongbing disease. *B. californicus*, *B. yothersi*, and *B. obovatus* are all present in Florida [@Childers2003; @Akyazi2017], and are difficult to identify by non-experts, or without advanced methodologies. DNA barcoding [@Armstrong2005] or a similarly simple and accurate method for identification of these mite complexes is vital to identify mite populations which need to be monitored or controlled. By doing so, we can determine the risk OFV-Orc represents for the native plants, agriculture and the ornamental/landscaping industries of Florida and the surrounding regions.

@@

### Figure captions
`r fig_nums(name = "fig_1", caption = "Variety of symptoms expressed by *Liriope* spp. infected with orchid fleck virus (OFV): (a) ringspot symptoms on *Liriope muscari* cv. 'Gigantea' (b-c) Details of ringspot symptoms on *L. muscari* cv. 'Gigantea' (d) rust colored spots on *Ophiopogon intermedius*")`

`r fig_nums(name = "fig_2", caption = "Symptoms expressed by *Aspidistra elatior* infected with OFV: (a) Detail of leaf chlorosis (b) Chlorosis appears similar to sun damage (c-d) Chlorotic ringspot may indicate early symptoms of OFV")`

>>>>

We describe the first detection of orchid fleck virus (OFV) infecting three unreported hosts: *Liriope muscari*, cv. 'Gigantea' (Decaisne) Bailey, *Ophiopogon intermedius* Don and *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) in Leon and Alachua Counties, FL. Orchid-infecting subgroup (Orc) of OFV infects over 50 plant species belonging to the plant families Orchidaceae, Asparagaceae (Nolinoidaea), and causes citrus leprosis disease in *Citrus* (Rutaceae). The only known vectors of OFV-Orc are the flat mites, *Brevipalpus californicus* (Banks) *sensu lato* (Trombidiformes: Tenuipalpidae). Florida has various plants in the landscape which *Brevipalpus* spp. feed on, which are susceptible to infection by OFV-Orc. Chlorotic ringspots and flecking were seen affecting Liriopogons (*Liriope* and *Ophiopogon* spp.) in Leon County, FL. Nearby *A. elatior* also appeared chlorotic. Local diagnostics returned negative for common plant pathogens, therefore new samples were sent to the Florida Department of Agriculture and Consumer Services (FDACS) and USDA-ARS for identification. Two orchid-infecting strains of OFV were detected via combinations of conventional RT-PCR, RT-qPCR, Sanger sequencing and High Throughput Sequencing (HTS). Amplicons shared 98\% nucleotide identity with OFV-Orc1 and OFV-Orc2 RNA2 genome sequences available in NCBI GenBank. Coinfections were detected in each county, but single strains of OFV-Orc were detected in *L. muscari* (Alachua, OFV-Orc2) and *A. elatior* (Leon, OFV-Orc1). Three potential mite vectors were identified via cryo-scanning electron microscopy (Cryo-SEM): *Brevipalpus californicus* (Banks) sensu lato, *B. obovatus* Donnadieu, and *B. confusus* Baker. In conclusion, OFV orchid strains are present in northern Florida, representing a risk for susceptible plants in the southeastern US.

@@

Se describe la primera detección del virus de orchid fleck virus (OFV), infectando a tres huéspedes no reportados: *Liriope muscari*, cv. ‘Gigantea’ (Decaisne) Bailey, *Ophiopogon intermedius* Don and *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) para los condados de Leon y Alachua, FL. Los subgrupos de OFV que infectan a las orquídeas (Orc) puedan infectar más de 50 especies de plantas pertenecientes a las familias Orchidaceae, Asparagaceae (Nolinoidaea), e infecta *Citrus* (Rutaceae) cómo la enfermedad de la leprosis de los cítricos. Los únicos vectores de OFV-Orc son los ácaros planos *Brevipalpus californicus* (Banks) *sensu lato* (Trombidiformes: Tenuipalpidae). La Florida tiene varias plantas en el campo ornamental, por lo cúales las especias de *Brevipalpus* se puedan alimentarse, y esas plantas son susceptibles a las infecciones de OFV-Orc. Se observaron manchas anulares cloróticas y salpicaduras en las hojas de las Serpentinas (*Liriope* y *Ophiopogon* spp.) y también se vieron hojas cloróticas en el *A. elatior* adyacente, situado en el condando de Leon, FL. Los diagnósticos del laboratorio local fueron negativos para los patógenos comunes, por lo tanto, se enviaron nuevos ejemplares al Florida Department of Agriculture and Consumer Services (FDACS) y el USDA-ARS para su identificación. Se detectaron dos cepas del OFV mediante la combinación de RT-PCR convencional, RT-qPCR, secuenciación de Sanger y secuenciación de alto rendimiento. Las ampliaciones compartían una identidad de nucleótidos del 98% con las secuencias del genoma de ARN2 de OFV-Orc1 y OFV-Orc2 disponibles en el NCBI GenBank. Se detectaron coinfecciones del virus en cada condado, pero se detectaron cepas únicas de OFV-Orc en *L. muscari* (Alachua , OFV-Orc2) y *A. elatior* (Leon , OFV-Orc1). Se identificaron tres ácaros mediante microscopía electrónica de barrido criogénico (Cryo-SEM) con potencial de ser vectores: *Brevipalpus californicus* (Banks) *sensu lato*, *B. obovatus* Donnadieu y *B. confusus* Baker. En conclusión, OFV está presente en el norte de Florida, lo que representa un riesgo para las plantas susceptibles en el sureste de Estados Unidos.

@@

\pagebreak
Orchid fleck virus (OFV), is the type member for the genus *Dichorhavirus*, family *Rhabdoviridae*. The virus is a bacilliform, nuclear rhabdovirus composed of two segments of single-stranded, negative-sense RNA which infects plants [@Amarasinghe2019; @Dietzgen2014; @Walker2018]. Flat mites from the genus *Brevipalpus* (Trombidiformes: Tenuipalpidae) are the only which have been experimentally shown to transmit dichorhaviruses [@Maeda1998]. Plants infected with OFV exhibit chlorotic and necrotic flecks on their leaves[@Kubo2009a; @Kubo2009b; @Dietzgen2018b]. The virus was first described as infecting *Cymbidium* orchids in Japan [@Doi1977]. There have been reports of OFV and OFV-like rhabdoviruses infecting orchids in Asia, Africa, North America, South America, Europe, and Oceania. The prevalence of OFV and its mite vector is thought to be associated with the movement of infected orchids [@Dietzgen2018a]. More than fifty species of Orchidaceae [@Kitajima2010; @Peng2013] can naturally become infected with OFV, as well as some Asparagaceae (Nolinoidaea) [@Mei2016; @Dietzgen2018b], and Rutaceae, where infection causes citrus leprosis-like symptoms [@Cook2019; @Velarde2021; @Roy2015; @Roy2020]. Mechanical transmission of OFV is possible under laboratory conditions to the plant families Chenopodiaceae, Aizoaceae, Fabaceae, and Solanaceae [@Chang1976; @Kondo2003; @Peng2013].

@@

During June 2020, chlorotic flecks and ringspot patterns of unknown etiology were observed on Giant Lilyturf *Liriope* spp., cv. 'Gigantea' in a landscape of Leon County, Florida (`r citef("fig_1")`). *Liriope* belong to a group of plants in the family Asparagaceae, subfamily Nolinoidaea, comprised of grass-like monocotyledonous liliod plants native to southeastern Asia [@Chase2009; @Meng2021]. *Liriope* and the closely related *Ophiopogon* (Asparagaceae: Nolinoidaea) are considered the most important ground cover plant in the southeastern United States [@Mcharo2003]. Viral infections of suspected leaf samples were initially tested at the Plant Disease Diagnostic Clinic at the North Florida Research and Education Center (NFREC) in Quincy, FL. All the samples were tested with one step conventional RT-PCR, and were found negative for begomovirus, carlavirus, potyvirus, tospovirus, cucumber mosaic virus and tobacco mosaic virus. As initial diagnostics were inconclusive, samples were taken of putatively infected plants with ringspot symptoms during July and August of 2020. Leaves were taken from *Liriope* spp. and *Ophiopogon* spp., as well as the *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea), nearby, which appeared sickly and chlorotic (`r citef("fig_2")`). Plant materials were sent to the Florida Department of Agriculture and Consumer Services (FDACS) for identification. The FDACS determined that the pathogen was OFV using previously published primers and methods to conduct RT-PCR and Sanger sequencing [@Kubo2009a; @Kubo2009b; @RamosGonzalez2015].
The identity of the virus was verified as OFV Orchid strain 1, (OFV-Orc1), following the methods described in @Kondo2017. Nucleotide sequencing shared 98\% nucleotide identity with the OFV-isolates So (Accession No. AB244418) and Br (Accession No. MK522807), which belong to orchid subgroup I [@Kondo2006; @Kondo2017]. These samples from FDACS were subsequently retested by the USDA-APHIS-PPQ S&T Beltsville laboratory, in conjunction with tests of fresh samples from both Alachua and Leon counties. The USDA used RT-PCR, RT-qPCR, and High Throughput Sequencing (HTS) to reconfirm the presence of OFV. Conventional RT-PCR with Generic R2-Dicho-GF and R2-Dicho-GR primers amplifed ~800 nt of L-gene (RNA2) amplicon [@Roy2020], and OFV-Orc1 and OFV-Orc2 were detected in both *O. intermedius* and *A. elatior* from Leon County [@Kondo2017] (GenBank Accession Numbers: MZ852004, MZ852005 MZ852006, and MZ852007). 99% nucleotide sequence identity is shared between OFV-Orc1 and OFV-Orc2 for the RNA2 genome, whereas 90\% nt sequence identity was found between these two reassortment strains. The presence of OFV-Orc1 and OFV-Orc2 in Leon and Alachua counties was reaffirmed with HTS data (`r citet("tab_1")`): Analysis of HTS data from Leon County found that the symptomatic *L. muscari* were coinfected with both OFV-Orc1 and OFV-Orc2, while the symptomatic *A. elatior* were solely infected with OFV-Orc1. Sequence data of symptomatic *L. muscari* from Alachua County revealed infections with OFV-Orc2 (GenBank Accession MZ852006). After the initial identification by FDACS of OFV-Orc, mite samples were collected from symptomatic Asparagaceae in Leon County. Most mites collected were Tenuipalpid mites (flat mites or false spider mites), a pest of ornamental plants, some of which are known to act as vectors for plant viruses [@Childers2003; @Childers2011].

@@

Mite taxonomy is complicated by cryptic species complexes which occur in many plant-feeding groups of the Acari [@Arthur2011; @Skoracka2010; @Skoracka2013; @Umina1999], including tenuipalpid mites from the genus *Brevipalpus* [@Navia2013]. The commonly used phase-contrast microscopy is insufficient to detect some diagnostic characters for separation of cryptic species, instead best practices recommend the combination of Differential Interference Contrast (DIC) Microscopy and Scanning Electron Microscopy along with molecular methods to separate cryptic species [@Beard2015]. The flat mites collected were initially suspected to belong to *B. californicus* after inspection with phase contrast microscopy. Subsequent observation via DIC microscopy at FDACS agreed with this tentative identification. Unfortunately, the *B. californicus* s.l. species group, *sensu* @Baker1987 is suspected to contain cryptic species [@Childers2011; @Rodrigues2013]. New mite samples were collected from symptomatic liriopogons and *A. elatior* in Leon County and sent to USDA-ARS's Electron and Confocal Microscopy Unit for analysis. Three mite species were recovered and examined under cryo-scanning electron microscopy (Cryo-SEM): *B. californicus* s.l. (`r citef("fig_3")`), *B. obovatus* Donnadieu and *B. confusus* Baker. The recent report of OFV in the US is thought to be @Ko1985 which describes nuclear inclusions caused by an undescribed bacilliform rhabdovirus in *Brassia* orchids. The significance of this report is their description of the spoke-wheel configurations of the viral particles [@Ko1985], a sign typically associated with OFV infection [@Chang1976]. Unfortunately, this article made no mention of mites or further investigations of the virus. The first report of OFV in the continental US was @Bratsch2015, who confirmed the presence of OFV in *Phalaenopsis* hybrids using Transmission Electron Microscopy of ultrathin sections of plant tissue as well as molecular sequence analysis. They also discuss the association of OFV with *Brevipalpus* mites, but the authors did not make a conclusive species identification beyond suggesting that the mite vector belonged to the *B. californicus* group, referring to @Kondo2003's publication [@Bratsch2015]. Later reports of OFV described OFV infecting a previously undescribed Nolinoidaea hosts in Australia [@Dietzgen2018b; @Mei2016], including *Liriope spicata* (Thunb.) Lour, a different species of liriopogon than those identified from the Florida sites. We are not aware of any reports of OFV infecting liriopogons, *A. elatior* nor other Nolinoidaea in the US. Although @Zheng2013 had mentioned an association between *B. californicus* and *A. elatior*, they never reported symptoms of OFV-Orc in this plant. We believe that our findings indicate the first report of OFV-Orc infecting ornamental Nolinoidaea in Florida, and possibly the US. This publication also marks the first reports of *A. elatior* and *Ophiopogon* spp. as natural hosts of OFV-Orc. There are two orchid strains of OFV (OFV-Orc1 and OFV-Orc2), and two citrus strains (OFV-Cit1 and OFV-Cit2) [@Roy2020; @Beltran-Beltran2020]. The OFV strains detected in Florida are identical in genome sequence to the orchid strains of OFV infecting citrus in Hawaii, Mexico, Colombia, and South Africa [@Roy2020; @Beltran-Beltran2020]. Both OFV-Orc1 and OFV-Orc2 infect citrus [@Roy2020], but none of the citrus  strains have been reported from any orchid species. The *Brevipalpus* mites collected from liriopogons and *A. elatior* in Leon County were abundant on OFV-infected plants very near to citrus trees, some plants even surrounding the trunk. *B. californicus* s. l. has been reported as a pest of citrus [@Childers2003] and are often collected from citrus fruits [@Vacante2016; @Vacante2010; @Baker1987; @Baker1949]. The proximity of these mite vectors to citrus raises the question: why these trees are not currently infected with OFV-Orc? It is important to note the uncertainty surrounding the vector for OFV-Orc. There are three mite species which have been recovered from OFV-Orc infected plants: *B. obovatus*, and *B. confusus* and *B. californicus* s.l., but only *B. californicus* has been described as a vector of OFV. Even so, the *B. californicus* which we find on liriopogons and *A. elatior* may not be the same cryptic species as those found on citrus. Transmission of OFV from populations of *B. californicus* liriopogon/*A. elatior* to citrus may be limited by host preferences, vectorial capacity, viral propagation/circulation in the vector, viral acquisition times, or feeding times required for transmission to citrus. Even so, these types of questions require future study to determine the potential of nolinoidaea to citrus transmission. Best practices for integrated pest management have not been created for controlling *Brevipalpus* mites on these ornamentals, but methods designed to control *Brevipalpus* in other systems may be applicable. The most common method used to control *Bervipalpus* are synthetic acaricides [@Andrade2010; @Andrade2019]. Unfortunately, some acaricides and their residues can harm beneficial predatory mites as well [@Fernandez2017], even at low doses [@Havasi2021], and mixing different chemistries can be detrimental for mite control [@Vechia2018]. In addition, pesticide resistance has been reported in various *Brevipalpus* populations [@Rocha2021; @Alves2000; @Campos2002; @Omoto2000], due to exposure to pesticides used to control other arthropod pests [@Vechia2021]. In addition, predatory mites [@Argolo2020; @Chen2006], entomopathogenic fungi [@RossiZalaf2008; @Magalhaes2005; @Pena2015; @Revynthi2019] have shown promise for controlling other *Brevipalpus* mites. Moreover, it is often possible to integrate different control techniques for improved management, such as combining predatory mites with compatible acaricides and entomopathogenic fungi [@Reddy2001; @Midthassel2016; @Andrade2019]. In conclusion, detecting OFV in Florida represents a concern for horticulturists who grow orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagaceae species which are commonly used in landscaping. Florida is also home to a plethora of native and naturalized orchid species, many of which are threatened, including cultivated *Vanilla* in southern Florida [@Chambers2019] and the famous Ghost Orchid, [*Dendrophylax lindenii* (Lindl.) Benth. ex Rolfe]. Citrus leprosis was present in Florida during the 1860’s and almost eradicated by the mid-1960s [@Childers2003; @Knorr1968a; @Knorr1968b]. An examination of herbarium specimens of Florida citrus found that this historical virus, Citrus leprosis dichorhavirus-N0, is distantly related to the modern isolates of OFV [@Hartung2015; @Kitajima2011a; @Roy2020]. The recent detection of OFV-Orc1 in South Africa [@Cook2019] in *C. sinensis* (Navel and Valencia orange) and OFV-Orc2 in Hawaii [@Velarde2021] in *C. reticulata* (mandarin) and *C. jambhiri* (rough lemon) associated with leprosis-like symptoms highlights the potential threat of different isolates of OFV on citrus, which will be a definite concern to the US multi-billion-dollar citrus industry already impacted by the Huanglongbing disease. *B. californicus*, *B. yothersi*, and *B. obovatus* are all present in Florida [@Childers2003; @Akyazi2017], and are difficult to identify by non-experts, or without advanced methodologies. DNA barcoding [@Armstrong2005] or a similarly simple and accurate method for identification of these mite complexes is vital to identify mite populations which need to be monitored or controlled. By doing so, we can determine the risk OFV-Orc represents for the native plants, agriculture and the ornamental/landscaping industries of Florida and the surrounding regions.

@@

### Figure captions
`r fig_nums(name = "fig_1", caption = "Variety of symptoms seen on *Liriope* spp. infected with orchid fleck virus (OFV): (a) symptoms on *Liriope muscari* cv. 'Gigantea' (b-c) Details of symptoms on *L. muscari* cv. 'Gigantea' (d) rust colored spots on *Ophiopogon intermedius*")`

`r fig_nums(name = "fig_2", caption = "Symptoms seen on *Aspidistra elatior* infected with OFV: (a) Detail of leaf chlorosis (b) Chlorosis appears similar to sun damage (c-d) Chlorotic flecks may indicate early symptoms of OFV")`
